The Categorical Structure of Optics

### 2018-01-25 paper.tex
Typo
Reorder deckchairs
Updates

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\PreOptic}{\mathbf{PreOptic}}
\newcommand{\Optic}{\mathbf{Optic}}
\newcommand{\SemiOptic}{\mathbf{SemiOptic}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Prism}{\mathbf{Prism}}
\newcommand{\Hask}{\mathbf{Hask}}
\newcommand{\Endo}{\mathbf{Endo}}
\newcommand{\Strong}{\mathbf{Strong}}
\newcommand{\Tamb}{\mathbf{Tamb}}
\newcommand{\Point}{\mathbf{Point}}
\newcommand{\CoPoint}{\mathbf{CoPoint}}
\newcommand{\Traversable}{\mathbf{Traversable}}

@@

\title{The Categorical Structure of Profunctor Optics}
\author{Mitchell Riley}
\begin{document}
\maketitle

@@

\todo{TODO: Think about size issues, we are taking some very big coends...}

@@

\todo{ We elide boring uses of the structure maps so we can stay sane...}

\begin{proposition}
Let $\C$ be a category and $(\M, \otimes, I)$ a monoidal category that acts on $\C$. There is a category $\PreOptic_\M$ that has the same objects as $\C \times \C^\op$, with homsets given by
\begin{align*}
\PreOptic_\M((S, S'), (A, A')) = \int^{M \in \M} \C(S, M A) \times \C(M A', S')
\end{align*}
\end{proposition}
To distinguish these from morphisms in $\C \times \C^\op$, we will write preoptics with a crossed arrow: $p : (S, S') \hto (A, A')$. If $(l,r) : (S \to M A, M A' \to S')$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for their image in $\PreOptic_\M((S, S'), (A, A'))$, and say that $M$ is the complement for this representative.
\begin{proof}
In \cite[Section 6]{Doubles} this is proven abstractly, by exhibiting this category as the Kleisli category for a monad in the bicategory $\Prof$. Here we give a direct proof.

The composition map
\begin{align*}
\int^{M \in \M} \left(\C(S, M A) \times \C(M A', S')\right) \times \int^{M \in \M} \left( \C(R, M S) \times \C(M S', R')\right) \to \int^{M \in \M} \C(R, M A) \times \C(M A', R')
\end{align*}
of preoptics is induced by the following morphisms for every $M, N \in \M$:
\begin{align*}
&\C(S, M A) \times \C(M A', S') \times \C(R, N S) \times \C(N S', R')\\
\to \,& \C(NS, NM A) \times \C(NM A', NS') \times \C(R, N S) \times \C(N S', R') && \text{(functoriality of action of $N$)} \\
\to \,& \C(R, N M A) \times \C(N M A', R') && \text{(composition)} \\
\to \,& \C(R, (N \otimes M) A) \times \C((N \otimes M) A', R') && \text{(composition with $\mu_{M,N,A}$)} \\
\to \,&\int^{M \in \M} \C(R, M A) \times \C(M A', R') && \text{($\copr_{N \otimes M}$)}
\end{align*}
In other words, let $\langle l, r \rangle : (R, R') \hto (S, S')$ and $\langle l', r' \rangle : (S, S') \hto (A, A')$ with $M$ the complement for $\langle l, r \rangle$. The composite is then: $\langle l', r' \rangle \circ \langle l, r \rangle = \langle \mu_{M,N,A}(M l')l, r(Mr')\mu^{-1}_{M,N,A'} \rangle$.

The identity morphism $(S, S') \hto (S, S')$ is given by $\langle \epsilon_S, \epsilon_{S'}^{-1} \rangle$. \todo{which will later be described as $(\id_S, \id_{S'})_*$}

Associativity of composition can be checked equationally. Given $(l_1, r_1) : (R \to MS, MS' \to R')$ and $(l_2, r_2) : (S \to NA, NA' \to S')$ and $(l_3, r_3) : (A \to PB, PB' \to S')$:
\begin{align*}
(\langle l_3, r_3 \rangle \circ \langle l_2, r_2 \rangle) \circ \langle l_1, r_1 \rangle 
&= \langle (N l_3)l_2, r_2(Nr_3) \rangle \circ \langle l_1, r_1 \rangle \\
&= \langle M((N l_3)l_2)l_1, r_1M(r_2(Nr_3)) \rangle \\
&= \langle (M N l_3)(M l_2)l_1, r_1(M r_2)(MNr_3)) \rangle \\
&= \langle l_3, r_3 \rangle \circ (\langle (M l_2)l_1, r_1(Mr_2) \rangle) \\
&= \langle l_3, r_3 \rangle \circ (\langle l_2, r_2 \rangle \circ \langle l_1, r_1 \rangle)
\end{align*}

\todo{To justify this kind of equational reasoning: We use that products in $\Set$ preserve colimits, so the coends can be floated to the outside and we may pick representatives for all 3 preoptics simultaneously. In the middle we include into the coend then immediately pick a representative, so by the universal property of coends we may as well use the representative we already have. It seems like this would also work in an enriched setting (if you really wanted it to)}

Finally, we have the unit laws. For $(l, r) : (S \to MA, MA' \to S')$ we calculate:
\begin{align*}
\id_{A, A'} \circ \langle l, r\rangle 
&= \langle \epsilon_A, \epsilon_{A'}^{-1} \rangle \circ \langle l, r\rangle \\
&= \langle (M\epsilon_A) l, r (M\epsilon_{A'}^{-1})\rangle \\
&= \langle (\lambda_M A) l, r (\lambda_M^{-1} A')\rangle \\
&= \langle l, r (\lambda_M^{-1} A') (\lambda_M A')\rangle \\
&= \langle l, r \rangle  \\
\langle l, r \rangle \circ \id_{S, S'} 
&= \langle l, r \rangle \circ \langle \epsilon_S, \epsilon_{S'}^{-1} \rangle  \\
&= \langle (Il)\epsilon_S, \epsilon_{S'}^{-1} (Ir) \rangle \\
&= \langle (\lambda_M S)l, r (\lambda_{S'}^{-1} S') \rangle \\
&= \langle l, r (\lambda_{S'}^{-1} S')(\lambda_M S') \rangle \\
&= \langle l, r \rangle
\end{align*}
Where in both cases we have used the coend relations to move $\lambda_M$ to the other side.
\end{proof}

In \cite[Section 6]{Doubles} a very similar category was considered. The \emph{double} $\Double_l \C$ of a monoidal category $(\C, \otimes, I)$ is \todo{the opposite category of} a special case of the above, where $\C$ is allowed to act on itself via left tensor. We will denote this case by $\PreOptic_\otimes$.

\begin{proposition}
For any $\C$ and $\M$, there is a functor $(-, -)_* : \C \times \C^\op \to \PreOptic_\M$, which is given on objects by $(S, S')_* = (S, S')$ and for a morphism $(f, g) : (S, S') \to (A, A')$ by $(f, g)_* = \langle \epsilon_A f, g \epsilon_{A'}^{-1} \rangle$.
\end{proposition}
\begin{proof}
This preserves identities, as the identity on an object $(S, S')$ in $\PreOptic_\M$ is defined to be exactly $\langle \epsilon_S, \epsilon_{S'}^{-1} \rangle$.

To check functoriality, suppose we have $(f, g) : (S, S') \to (A, A')$ and $(f', g') : (A, A') \to (B, B')$ in $\C \times \C^\op$. Then:
\begin{align*}
(f', g')_* \circ (f, g)_* 
&= \langle \epsilon_B f', g' \epsilon_{B'}^{-1} \rangle \circ \langle \epsilon_A f, g \epsilon_{A'}^{-1} \rangle \\
&= \langle (I(\epsilon_B f'))\epsilon_A f, g \epsilon_{A'}^{-1} (I(g' \epsilon_{B'}^{-1}))\rangle && \text{(By definition of $\circ$)}\\
&= \langle (I\epsilon_B) (I f')\epsilon_A f, g \epsilon_{A'}^{-1} (I g')(I\epsilon_{B'}^{-1})\rangle && \text{(Functoriality of action of $I$)}\\
&= \langle (I\epsilon_B) \epsilon_B f' f, g g' \epsilon_{B'}^{-1} (I\epsilon_{B'}^{-1})\rangle && \text{(Naturality of $\epsilon$)}\\
&= \langle (\lambda_I B) \epsilon_B f' f, g g' \epsilon_{B'}^{-1} (\lambda_I^{-1}B')\rangle && \text{(Unitality of action)} \\
&= \langle \epsilon_B f' f, g g' \epsilon_{B'}^{-1} (\lambda_I^{-1}B') (\lambda_I B') \rangle && \text{(Coend relation)}  \\
&= \langle \epsilon_B f'f, g g' \epsilon_{B'}^{-1} \rangle \\
&= (f'f, gg')_*
\end{align*}
\end{proof}
\todo{Is this functor monoidal?}

\todo{Is this functor faithful? What if $I$ is terminal? No! Take $S = \emptyset$}

\todo{This doesn't give us the structure of a `framing', we can only lift vertical arrows to horizontal arrows facing one direction.}

\begin{theorem}
If $(\C, \otimes, I)$ is a symmetric monoidal category, then $\PreOptic_\otimes$ is also symmetric monoidal. \todo{braided?}
\end{theorem}
\begin{proof}
On objects we define $(S, S') \otimes (T, T') := (S \otimes T, S' \otimes T')$. The unit is given by the object $(I, I)$. The action of $\otimes$ on morphisms is induced by
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
\to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
\to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
\to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}
\end{align*}
yielding a function $\PreOptic_\otimes((S, S'), (A, A')) \times \PreOptic_\otimes((T, T'), (B, B')) \to \PreOptic_\otimes((S \otimes T, S' \otimes T'), (A \otimes B, A' \otimes B'))$.

Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
\begin{align*}
\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
\end{align*}
For future convenience, note that 
\begin{align*}
\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
&= \langle l \otimes T, r \otimes T' \rangle
\end{align*}
Now we check functoriality. Suppose we have:
\begin{align*}
\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
\end{align*}
with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
\begin{align*}
&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
\end{align*}

@@

Note that because $(S, S')_* = (S, S')$, required equations for $(-, -)_*$ to be a monoidal functor hold by definition (although we don't yet know that $\PreOptic_\otimes$ is monoidal). The pentagon and triangle axioms then hold in $\PreOptic_\otimes$, as they are the image of the same axioms in $\C \times \C^\op$ under $(-, -)_*$. The only remaining piece to verify is that these structure maps are natural in $\PreOptic_\otimes$.

@@

There is some further useful structure in $\PreOptic_\otimes$. The set of costates $(S, S') \hto (I, I)$ is isomorphic to $\C(S, S')$:
\begin{align*}
\PreOptic_\otimes((S, S'), (I, I)) 
&= \int^{M \in \C} \C(S, M \otimes I) \times \C(M \otimes I, S') \\
&\cong \int^{M \in \C} \C(S, M) \times \C(M, S') \\
&\cong \C(S, S')
\end{align*}
The identity morphism $\C(S, S)$ gives us a preoptic $\varepsilon : (S, S) \to (I, I)$ that we call the \emph{counit}.

States are not as easy to describe in general.

>>>>

% Categories
\newcommand{\Set}{\mathbf{Set}}
\newcommand{\Cat}{\mathbf{Cat}}
\newcommand{\Prof}{\mathbf{Prof}}
\newcommand{\MonCat}{\mathbf{MonCat}}
\newcommand{\LaxMonCat}{\mathbf{LaxMonCat}}
\newcommand{\SymmMonCat}{\mathbf{SymmMonCat}}

\newcommand{\Act}{\mathbf{Act}}
\newcommand{\Optic}{\mathbf{Optic}}
\newcommand{\Lawful}{\mathbf{Lawful}}
\newcommand{\SemiOptic}{\mathbf{SemiOptic}}
\newcommand{\Lens}{\mathbf{Lens}}
\newcommand{\Prism}{\mathbf{Prism}}
\newcommand{\Hask}{\mathbf{Hask}}
\newcommand{\Endo}{\mathbf{Endo}}
\newcommand{\Strong}{\mathbf{Strong}}
\newcommand{\Tamb}{\mathbf{Tamb}}
\newcommand{\Point}{\mathbf{Point}}
\newcommand{\CoPoint}{\mathbf{CoPoint}}
\newcommand{\Traversable}{\mathbf{Traversable}}

@@

\title{The Categorical Structure of Optics}
\author{Mitchell Riley}
\begin{document}
\maketitle

@@

@@

We begin by constructing the category of optics in a symmetric monoidal category $\C$. This category was first defined in \cite[Section 6]{Doubles} to investigate the relationship between Tambara modules and the center of a monoidal category. There it is called the \emph{double} $\Double_l \C$ of a monoidal category, and our definition is identical, other than flipping the direction of the arrows to match the existing literature on lenses.

\begin{remark}
\todo{I hope I can get away with this...} In what follows, we use the standard cheat of omitting associativity isomorphisms
\end{remark}

Let $\C$ be a symmetric monoidal category, and suppose we have two pairs of objects $(S, S')$ and $(A, A')$. An \emph{optic} $p : (S, S') \hto (A, A')$ is an element of the set
\begin{align*}
\Optic_\C((S, S'), (A, A')) := \int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')
\end{align*}
If $l : S \to M \otimes A$ and $r : M \otimes A' \to S'$, we write $\langle l, r \rangle : (S, S') \hto (A, A')$ for the image of this pair in $\Optic_\C((S, S'), (A, A'))$, and say that $M$ is the complement for this representative.

\begin{remark}
Optics will always be written with a crossed arrow $\hto$, to distinguish them from morphisms of $\C$.
\end{remark}

Elements of $\Optic_\C((S, S'), (A, A'))$ have a suggestive interpretation as string diagrams with a ``hole'' missing. We draw the pair $\langle l, r \rangle$ as
\begin{center}
\input{diagrams/generic-optic.tikz}
\end{center}
using the \todo{strange} convention of reading string diagrams from bottom to top.

Two optics can be composed as follows: \todo{I give possibly too much detail here.} We wish to construct a map
\begin{align*}
&\left(\int^{M \in \C} \C(S, M \otimes A) \times \C(M \otimes A', S')\right) \times \left(\int^{M \in \C} \C(R, M \otimes S) \times \C(M \otimes S', R')\right) \\ &
 \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
The product in $\Set$ preserves colimits, so in particular coends. Using this fact, and the Fubini theorem for coends \cite[Remark 1.9]{CoendCofriend}, the domain is isomorphic to
\begin{align*}
\int^{(M, N) \in \C \times \C} \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R').
\end{align*}
So by the universal property of coends, it suffices to construct natural maps 
\begin{align*}
& \C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') \\ &
 \quad \to \int^{M \in \C} \C(R, M \otimes A) \times \C(M \otimes A', R').
\end{align*}
for each $M, N \in \C$. For this we use the composites
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(R, N \otimes S) \times \C(N \otimes S', R')\\
\to \,& \C(N \otimes S, N \otimes M \otimes A) \times \C(N \otimes M \otimes A', N \otimes S') \times \C(R, N \otimes S) \times \C(N \otimes S', R') && \text{(functoriality of $N \otimes  -$)} \\
\to \,& \C(R, N \otimes  M \otimes A) \times \C(N \otimes M \otimes A', R') && \text{(composition in $\C$)} \\
\to \,&\int^{M \in \M} \C(R, M \otimes A) \times \C(M \otimes A', R') && \text{($\copr_{N \otimes M}$)}
\end{align*}
In other words, let $\langle l, r \rangle : (R, R') \hto (S, S')$ and $\langle l', r' \rangle : (S, S') \hto (A, A')$ with $M$ the complement for $\langle l, r \rangle$. The composite is then: $\langle l', r' \rangle \circ \langle l, r \rangle = \langle (M l')l, r(Mr')\rangle$.

Graphically, this is placing the second optic into the hole of the first:
\begin{center}
\todo{todo}
\end{center}

The identity optic $(S, S') \hto (S, S')$ is given by $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$, where $\lambda_S : I \otimes S \to S$ is the left unitor for $S$.

\begin{proposition}
The above data form a category $\Optic_\C$.
\end{proposition}
\begin{proof}
In \cite[Section 6]{Doubles} this is proven abstractly, by exhibiting this category as the Kleisli category for a monad in the bicategory $\Prof$. Here we prove the properties directly.

We check associativity of composition equationally. As in the definition of composition, the Fubini theorem allows us to choose representatives for all three optics simultaneously. Suppose we have representatives of three optics $\langle l_1, r_1\rangle : (R, R) \hto (S, S')$ and $\langle l_2, r_2 \rangle : (S, S') \hto (A, A')$ and $\langle l_3, r_3 \rangle : (A, A') \hto (B, B')$, that have complements $M$, $N$ and $P$ respectively. Then:
\begin{align*}
(\langle l_3, r_3 \rangle \circ \langle l_2, r_2 \rangle) \circ \langle l_1, r_1 \rangle 
&= \langle (N \otimes l_3)l_2, r_2(N \otimes r_3) \rangle \circ \langle l_1, r_1 \rangle \\
&= \langle (M \otimes ((N \otimes l_3)l_2))l_1, r_1(M \otimes (r_2(N \otimes r_3))) \rangle \\
&= \langle (M \otimes N \otimes l_3)(M \otimes l_2)l_1, r_1(M \otimes r_2)(M \otimes N \otimes r_3)) \rangle \\
&= \langle l_3, r_3 \rangle \circ (\langle (M \otimes l_2)l_1, r_1(M \otimes r_2) \rangle) \\
&= \langle l_3, r_3 \rangle \circ (\langle l_2, r_2 \rangle \circ \langle l_1, r_1 \rangle)
\end{align*}
Graphically:
\begin{center}
\todo{todo}
\end{center}

For the unit laws, suppose we have $\langle l, r \rangle : (S, S') \hto (A, A')$ with representative $M$. We calculate:
\begin{align*}
\id_{A, A'} \circ \langle l, r\rangle 
&= \langle \lambda^{-1}_A, \lambda_{A'} \rangle \circ \langle l, r\rangle \\
&= \langle (M \otimes \lambda^{-1}_A) l, r (M\otimes  \lambda_{A'})\rangle \\
&= \langle (\lambda^{-1}_M \otimes  A) l, r (\lambda_M \otimes A')\rangle \\
&= \langle l, r (\lambda_M \otimes A') (\lambda^{-1}_M \otimes A')\rangle \\
&= \langle l, r \rangle  \\
\langle l, r \rangle \circ \id_{S, S'} 
&= \langle l, r \rangle \circ \langle \lambda^{-1}_S, \lambda_{S'}\rangle  \\
&= \langle (I \otimes l)\lambda^{-1}_S, \lambda_{S'} (I \otimes r) \rangle \\
&= \langle (\lambda^{-1}_M \otimes S)l, r (\lambda_{M} \otimes S') \rangle \\
&= \langle l, r (\lambda_{M} \otimes S')(\lambda^{-1}_M \otimes S') \rangle \\
&= \langle l, r \rangle
\end{align*}
Where in both cases we have used the naturality of $\lambda$ and coend relations to move $\lambda^{-1}_M$ to the right-hand side of the coend. Graphically:
\begin{center}
\todo{todo}
\end{center}

%<---
\end{proof}

\begin{remark}
Note that the definition of the $\Optic$ category involves a coend possibly indexed by a large category. To be maximally careful, we should only ever discuss $\Optic$ categories where we know that the coend exists by some other means, e.g., by exhibiting an isomorphism with some other set. For many, but not all, of the examples given later, we provide such an isomorphism.

\todo{How does the doubles paper get away with it?}
\end{remark}

\begin{proposition}
For any $\C$, there is a functor $(-, -)_* : \C \times \C^\op \to \Optic_\C$, which is given on objects by $(S, S')_* = (S, S')$ and for a morphism $(f, g) : (S, S') \to (A, A')$ by $(f, g)_* = \langle \lambda_A^{-1} f, g \lambda_{A'} \rangle$.
\end{proposition}
\begin{proof}
This preserves identities, as the identity on an object $(S, S')$ in $\Optic_\M$ is defined to be exactly $\langle \lambda^{-1}_S, \lambda_{S'} \rangle$.

To check functoriality, suppose we have $(f, g) : (S, S') \to (A, A')$ and $(f', g') : (A, A') \to (B, B')$ in $\C \times \C^\op$. Then:
\begin{align*}
(f', g')_* \circ (f, g)_* 
&= \langle \lambda^{-1}_B f', g' \lambda_{B'} \rangle \circ \langle \lambda^{-1}_A f, g \lambda_{A'} \rangle \\
&= \langle (I\otimes (\lambda^{-1}_B f'))\lambda^{-1}_A f, g \lambda_{A'} (I\otimes (g' \lambda_{B'}))\rangle && \text{(By definition of $\circ$)}\\
&= \langle (I \otimes \lambda^{-1}_B) (I \otimes f')\lambda^{-1}_A f, g \lambda_{A'} (I \otimes g')(I\otimes \lambda_{B'})\rangle && \text{(Functoriality of $I \otimes -$)}\\
&= \langle (I\otimes \lambda^{-1}_B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (I\otimes \lambda_{B'})\rangle && \text{(Naturality of $\lambda$)}\\
&= \langle (\lambda^{-1}_I \otimes B) \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B')\rangle && \text{(Unitality of action)} \\
&= \langle \lambda^{-1}_B f' f, g g' \lambda_{B'} (\lambda_I \otimes B') (\lambda^{-1}_I \otimes B') \rangle && \text{(Coend relation)}  \\
&= \langle \lambda^{-1}_B f'f, g g' \lambda_{B'} \rangle \\
&= (f'f, gg')_*
\end{align*}
Graphically, there is not much to do:
\begin{center}
\todo{todo}
\end{center}
\end{proof}

%\todo{This doesn't give us the structure of a `framing', we can only lift vertical arrows to horizontal arrows facing one direction.}

\begin{theorem}
$\Optic_\C$ is symmetric monoidal, where $(S, S') \otimes (T, T') = (S \otimes T, S' \otimes T')$, the unit is $(I, I)$, and with the action on morphisms induced by
\begin{align*}
&\C(S, M \otimes A) \times \C(M \otimes A', S') \times \C(T, N \otimes B) \times \C(N \otimes B', T')\\
\to \quad&\C(S \otimes T, M \otimes A \otimes N \otimes B) \times \C(M \otimes A' \otimes N \otimes B', S' \otimes T') && \text{(action of $\otimes$ on morphisms in $\C$)}\\
\to \quad&\C(S \otimes T, M \otimes N \otimes A \otimes B) \times \C(M \otimes N \otimes A' \otimes B', S' \otimes T') && \text{($\otimes$ in $\C$ is symmetric)}\\
\to \quad&\int^{M \in \C} \C(S \otimes T, M \otimes A \otimes B) \times \C(M \otimes A' \otimes B', S' \otimes T') && \text{(inclusion into coend)}
\end{align*}
\end{theorem}
\begin{proof}
Equationally, suppose we have representatives $(l, r) : (S \to M \otimes A, M \otimes A' \to S')$ and $(l', r') : (T \to N \otimes B, N \otimes B' \to T')$. Then:
\begin{align*}
\langle l, r \rangle \otimes \langle l', r' \rangle &:= \langle (M \otimes s_{A,N} \otimes B)(l \otimes l'), (r \otimes r')(M \otimes s_{A',N} \otimes B') \rangle
\end{align*}
For future convenience, note that 
\begin{align*}
\langle l, r \rangle \otimes (T, T') &= \langle (M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T') \rangle \\
&= \langle (\lambda^{-1}_M \otimes A \otimes T)(M \otimes s_{A,I} \otimes T)(l \otimes \lambda_T), (r \otimes \lambda^{-1}_{T'})(M \otimes s_{A',I}\otimes T')(\lambda_M \otimes A' \otimes T') \rangle \\
&= \langle l \otimes T, r \otimes T' \rangle
\end{align*}
Now we check functoriality. Suppose we have:
\begin{align*}
\langle l_1, r_1 \rangle : (S_1, S_1') &\hto (S_2, S_2') \\
\langle l_2, r_2 \rangle : (S_2, S_2') &\hto (S_3, S_3') \\
\langle p_1, q_1 \rangle : (T_1, T_1') &\hto (T_2, T_2') \\
\langle p_2, q_2 \rangle : (T_2, T_2') &\hto (T_3, T_3')
\end{align*}
with complements $M_1, M_2, N_1, N_2$ respectively. Then: \todo {(This is a mess. String diagram?)}
\begin{align*}
&(\langle l_2, r_2 \rangle \otimes \langle p_2, q_2 \rangle) \circ (\langle l_1, r_1 \rangle \otimes \langle p_1, q_1 \rangle) \\
= \; &\langle (M_2 \otimes s_{S_3,N_2} \otimes T_3)(l_2 \otimes p_2), (r_2 \otimes q_2)(M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
&\circ \langle (M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), (r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes N_1 \otimes l_2 \otimes p_2)(M_1 \otimes s_{S_2,N_1} \otimes T_2)(l_1 \otimes p_1), \\
&\;(r_1 \otimes q_1)(M_1 \otimes s_{S_2',N_1} \otimes T_2')(M_1 \otimes N_1 \otimes r_2 \otimes q_2)(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)(M_1 \otimes  l_2 \otimes N_1 \otimes p_2)(l_1 \otimes p_1), \\
&\;(r_1 \otimes q_1)(M_1 \otimes r_2 \otimes N_1 \otimes q_2)(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \langle (M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3,N_2} \otimes T_3)(M_1 \otimes s_{(M_2 \otimes S_3),N_1} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1), \\
&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes s_{(M_2 \otimes S_3'),N_1} \otimes T_3')(M_1 \otimes N_1 \otimes M_2 \otimes s_{S_3',N_2} \otimes T_3') \rangle \\
= \; & \text{\todo{(Add a $s_{M_2, N_1}$ to either side of the coend, then appeal to coherence)}} \\
= \; &\langle (M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S_3} \otimes T_3)((M_1 \otimes l_2)l_1\otimes (N_1 \otimes p_2)p_1) ,  \\
&\;(r_1(M_1 \otimes r_2) \otimes q_1(N_1 \otimes q_2))(M_1 \otimes M_2 \otimes s_{(N_1 \otimes N_2),S'_3} \otimes T'_3) \rangle \\
= \; &\langle (M_1 \otimes l_2)l_1, r_1(M_1 \otimes r_2) \rangle \otimes \langle (N_1 \otimes p_2)p_1, q_1(N_1 \otimes q_2) \rangle \\
= \; &(\langle l_2, r_2 \rangle  \circ \langle l_1, r_1 \rangle) \otimes (\langle p_2, q_2 \rangle \circ \langle p_1, q_1 \rangle)
\end{align*}

@@

Note that because $(S, S')_* = (S, S')$, required equations for $(-, -)_*$ to be a monoidal functor hold by definition (although we don't yet know that $\Optic_\otimes$ is monoidal). The pentagon and triangle axioms then hold in $\Optic_\otimes$, as they are the image of the same axioms in $\C \times \C^\op$ under $(-, -)_*$. The only remaining piece to verify is that these structure maps are natural in $\Optic_\otimes$.

@@

There is some further useful structure in $\Optic_\C$. 
\begin{proposition}
The set of costates $(S, S') \hto (I, I)$ is isomorphic to $\C(S, S')$.
\end{proposition}
\begin{proof}
\begin{align*}
\Optic_\otimes((S, S'), (I, I)) 
&= \int^{M \in \C} \C(S, M \otimes I) \times \C(M \otimes I, S') \\
&\cong \int^{M \in \C} \C(S, M) \times \C(M, S') \\
&\cong \C(S, S')
\end{align*}
\end{proof}

States are not as easy to describe in general:
